\chapter*{About This Monograph}
\addcontentsline{toc}{chapter}{About This Monograph}
\markboth{About This Monograph}{About This Monograph}

Downstream processing---the sequence of unit operations that turns a crude
biological harvest into a pure, formulated drug substance---is where much of the
cost, the risk, and the intellectual difficulty of a modern biomanufacturing
process resides. Of those unit operations, \emph{preparative chromatography} is
the workhorse. It is also, mathematically, the richest: a chromatographic column
is a distributed-parameter system in which convection, dispersion, and
nonlinear, multi-component adsorption interact, and whose behaviour swings by
orders of magnitude as a handful of mobile-phase variables are turned.

This monograph documents the scientific content of the \code{DownstreamDOE}
software package: a \emph{virtual laboratory} for downstream bioprocess
development. The package does two things. First, it provides mechanistic,
first-principles models of chromatography that act as a synthetic ``ground
truth''---an oracle that, when probed with a set of process conditions, returns
the chromatogram and the pool quality attributes that a real experiment would.
Second, it provides the statistical and machine-learning machinery of
\emph{Design of Experiments} (DoE) that one uses to interrogate such an oracle
efficiently: factorial designs, space-filling designs, latent-variable analysis,
and adaptive Bayesian optimisation.

The document is organised into two parts. \textbf{Part~I (Foundations)} is for the
reader who arrives strong in one of this book's two parent disciplines but new to
the other. Because the monograph fuses statistics with downstream separation
science, it is easy to assume of either field a fluency the reader may not yet
have; these two chapters supply it, each built from the ground up but curated to
exactly the ideas the later chapters use.

\begin{itemize}
  \item \textbf{Chapter~\ref{ch:stats}} introduces the statistical vocabulary---
  distributions, estimation, least squares and the analysis of variance,
  cross-validation, the bootstrap, the Bayesian update, and the crucial
  distinction between \emph{variability} and \emph{uncertainty}---that the design
  methods rest on.

  \item \textbf{Chapter~\ref{ch:sep}} introduces the separation science---unit
  operations, the molecules and their impurities, the adsorption isotherm, the
  column and its plates and resolution, mass transfer, the modulator idea that
  unifies the chromatography modes, and the basics of microbial fermentation.
\end{itemize}

\textbf{Part~II (The Virtual Laboratory)} is the technical core.

\begin{itemize}
  \item \textbf{Chapter~\ref{ch:models}} develops the chromatography models. We
  begin from the governing conservation laws, reduce them to the equations the
  package actually integrates, and then treat each adsorption mode---ion
  exchange, hydrophobic interaction, and reversed phase---in turn. For every
  model we give the full mathematical formulation, a careful description of each
  parameter and its units, the real-world separation problems the model
  represents, the assumptions and limitations that bound its validity, and the
  experiments and data one performs to estimate its parameters.

  \item \textbf{Chapter~\ref{ch:design}} develops the experimental-design
  methods. Each method---full factorial designs with response-surface analysis,
  Latin hypercube sampling, principal-component and partial-least-squares
  analysis, and Gaussian-process Bayesian optimisation---is presented with its
  theoretical underpinning, its mathematical formulation, and an honest
  accounting of where it is strong and where it fails.

  \item \textbf{Chapter~\ref{ch:ferment}} turns to a different kind of unit
  operation---a living \emph{milk fermentation}, the basis of yogurt and cultured
  dairy. Here the model is mechanistic but irreducibly stochastic, the only
  observable is an indirect pH curve, and the central design variable is
  \emph{combinatorial}: which of many candidate strains to blend. The chapter
  develops the fermentation model and its three layers of randomness, the
  real-world experiment, the \emph{covering-array} and related designs for
  screening large strain libraries, and the tree-ensemble methods used to analyse
  the resulting data.
\end{itemize}

Throughout, the mathematics is tied back to the code. Where a symbol in an
equation corresponds to an argument of a function or a field of a data
structure, we say so, in \code{typewriter} font. The intent is that a reader can
move from the page to the source and back without losing the thread.

A word on philosophy. The models in this package are deliberately the
\emph{simplest models that retain the right physics}. The linearised
Steric-Mass-Action isotherm, the lumped transport-dispersive band-broadening
term, the film model of ultrafiltration flux: each is a reduction of a more
complete theory, chosen because it captures the dependence of the response on
the critical process parameters faithfully while remaining cheap enough to
evaluate thousands of times inside a design loop. We have tried to be explicit
about every such reduction, because knowing the boundary of a model's validity
is the difference between a tool and a trap.

\vspace{1em}
{\itshape\noindent
The figures in this monograph are generated directly from the package source by
\code{doc/scripts/make\_figures.py} (and, for the foundations chapters,
\code{make\_foundations\_figures.py}); the chromatograms are real numerical
solutions of the models described in the text, and even the introductory figures
are computed from the small teaching module
\code{downstream\_doe.foundations}---not schematics.\par}
