\chapter*{Authors \& License}
\addcontentsline{toc}{chapter}{Authors \& License}
\markboth{Authors \& License}{Authors \& License}

This is an \emph{auto-generated} monograph. Its text, derivations, and chapter
structure were drafted by \textbf{Claude Code} (Anthropic) directly from the
mathematical \emph{formulation} encoded in the \code{DownstreamDOE} package---the
mechanistic models, their parameters and units, and the design methods---under the
direction of \textbf{Ali A. Eftekhari}, who conceived the virtual-laboratory idea,
supplied the underlying models and formulations, and edited the result. The prose
follows the code rather than the other way around, and every figure is a real
numerical solution produced by the package itself, not a schematic.

\begin{itemize}
  \item \textbf{Ali A. Eftekhari} --- concept, mechanistic formulations, package
  design, and editorial direction.
  \item \textbf{Claude Code} (Anthropic) --- drafting of the \code{DownstreamDOE}
  Python package and of this monograph from those formulations.
\end{itemize}

\begin{keybox}{MIT License}
\noindent Copyright \textcopyright{} 2026 Ali A. Eftekhari

\smallskip
\noindent Permission is hereby granted, free of charge, to any person obtaining a
copy of this software and associated documentation files (the ``Software''), to
deal in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell copies of
the Software, and to permit persons to whom the Software is furnished to do so,
subject to the following conditions:

\smallskip
\noindent The above copyright notice and this permission notice shall be included
in all copies or substantial portions of the Software.

\smallskip
\noindent THE SOFTWARE IS PROVIDED ``AS IS'', WITHOUT WARRANTY OF ANY KIND, EXPRESS
OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, FITNESS
FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR
COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN
ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH
THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
\end{keybox}
