\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
\markboth{Preface}{Preface}

Downstream processing---the sequence of unit operations that turns a crude
biological harvest into a pure, formulated drug substance---is where much of the
cost, the risk, and the intellectual difficulty of a modern biomanufacturing
process resides. Of those unit operations, \emph{preparative chromatography} is
the workhorse. It is also, mathematically, the richest: a chromatographic column
is a distributed-parameter system in which convection, dispersion, and
nonlinear, multi-component adsorption interact, and whose behaviour swings by
orders of magnitude as a handful of mobile-phase variables are turned.

This monograph documents the scientific content of the \code{DownstreamDOE}
software package: a \emph{virtual laboratory} for downstream bioprocess
development. The package does two things. First, it provides mechanistic,
first-principles models of chromatography that act as a synthetic ``ground
truth''---an oracle that, when probed with a set of process conditions, returns
the chromatogram and the pool quality attributes that a real experiment would.
Second, it provides the statistical and machine-learning machinery of
\emph{Design of Experiments} (DoE) that one uses to interrogate such an oracle
efficiently: factorial designs, space-filling designs, latent-variable analysis,
and adaptive Bayesian optimisation.

The document is organised into three chapters.

\begin{itemize}
  \item \textbf{Chapter~\ref{ch:models}} develops the chromatography models. We
  begin from the governing conservation laws, reduce them to the equations the
  package actually integrates, and then treat each adsorption mode---ion
  exchange, hydrophobic interaction, and reversed phase---in turn. For every
  model we give the full mathematical formulation, a careful description of each
  parameter and its units, the real-world separation problems the model
  represents, the assumptions and limitations that bound its validity, and the
  experiments and data one performs to estimate its parameters.

  \item \textbf{Chapter~\ref{ch:design}} develops the experimental-design
  methods. Each method---full factorial designs with response-surface analysis,
  Latin hypercube sampling, principal-component and partial-least-squares
  analysis, and Gaussian-process Bayesian optimisation---is presented with its
  theoretical underpinning, its mathematical formulation, and an honest
  accounting of where it is strong and where it fails.

  \item \textbf{Chapter~\ref{ch:ferment}} turns to a different kind of unit
  operation---a living \emph{milk fermentation}, the basis of yogurt and cultured
  dairy. Here the model is mechanistic but irreducibly stochastic, the only
  observable is an indirect pH curve, and the central design variable is
  \emph{combinatorial}: which of many candidate strains to blend. The chapter
  develops the fermentation model and its three layers of randomness, the
  real-world experiment, the \emph{covering-array} and related designs for
  screening large strain libraries, and the tree-ensemble methods used to analyse
  the resulting data.
\end{itemize}

Throughout, the mathematics is tied back to the code. Where a symbol in an
equation corresponds to an argument of a function or a field of a data
structure, we say so, in \code{typewriter} font. The intent is that a reader can
move from the page to the source and back without losing the thread.

A word on philosophy. The models in this package are deliberately the
\emph{simplest models that retain the right physics}. The linearised
Steric-Mass-Action isotherm, the lumped transport-dispersive band-broadening
term, the film model of ultrafiltration flux: each is a reduction of a more
complete theory, chosen because it captures the dependence of the response on
the critical process parameters faithfully while remaining cheap enough to
evaluate thousands of times inside a design loop. We have tried to be explicit
about every such reduction, because knowing the boundary of a model's validity
is the difference between a tool and a trap.

\vspace{1em}
\noindent\itshape
The figures in this monograph are generated directly from the package source by
\code{doc/scripts/make\_figures.py}; the chromatograms are real numerical
solutions of the models described in the text, not schematics.
